---
title: "Project"
output: html_notebook
---



```{r}
rm(list=ls())
ls()

setwd("F:/Senior-2-Spring/Big-Data/Project-Repo/Big-Data-CrimeAnalysis-R-Hadoop") 
```

load csv
```{r}
dataset <- read.csv("Major_Crime_Indicators.csv")

anyNA(dataset$mci_category)
anyNA(dataset$reportedday)
anyNA(dataset$reportedyear)
anyNA(dataset$reportedmonth)
anyNA(dataset$reportedhour)
anyNA(dataset$offence)
anyNA(dataset$Neighbourhood)
anyNA(dataset$premises_type)
anyNA(dataset$occurrenceday)
anyNA(dataset$occurrencehour)
anyNA(dataset$occurrencemonth)
anyNA(dataset$occurrenceyear)
anyNA(dataset$occurrencedate)

#=========================Filling missing value====================================
date_year<-c(dataset$occurrencedate[is.na(dataset$occurrenceyear)])
#filling missing data in occurrence year from the column occurrence date
dataset$occurrenceyear[is.na(dataset$occurrenceyear)]<-substr(date_year,1,4)
#checking that they are all filled
anyNA(dataset$occurrenceyear)
levels(factor(dataset$occurrenceyear))
date_day<-c(dataset$occurrencedate[is.na(dataset$occurrenceday)])

#filling missing data in occurrence day from the column occurrence date
dataset$occurrenceday[is.na(dataset$occurrenceday)]<-substr(date_day,9,10)
#checking that they are all filled
anyNA(dataset$occurrenceday)
levels(factor(dataset$occurrenceday))

date_month<-c(dataset$occurrencedate[dataset$occurrencemonth==""])

substr(date_month[1],6,7)
#filling missing data in occurrence day from the column occurrence date
dataset$occurrencemonth[dataset$occurrencemonth==""]<-substr(date_month,6,7)
levels(factor(dataset$occurrencemonth))
dataset$occurrencemonth[which(dataset$occurrencemonth=="January")] <- "1"
dataset$occurrencemonth[which(dataset$occurrencemonth=="February")] <- "2"
dataset$occurrencemonth[which(dataset$occurrencemonth=="March")] <- "3"
dataset$occurrencemonth[which(dataset$occurrencemonth=="April")] <- "4"
dataset$occurrencemonth[which(dataset$occurrencemonth=="May")] <- "5"
dataset$occurrencemonth[which(dataset$occurrencemonth=="June")] <- "6"
dataset$occurrencemonth[which(dataset$occurrencemonth=="July")] <- "7"
dataset$occurrencemonth[which(dataset$occurrencemonth=="August")] <- "8"
dataset$occurrencemonth[which(dataset$occurrencemonth=="September")] <- "9"
dataset$occurrencemonth[which(dataset$occurrencemonth=="October")] <- "10"
dataset$occurrencemonth[which(dataset$occurrencemonth=="November")] <- "11"
dataset$occurrencemonth[which(dataset$occurrencemonth=="December")] <- "12"
dataset$occurrencemonth <- as.numeric(dataset$occurrencemonth)
levels(factor(dataset$occurrencemonth))



week_day<-c(dataset$occurrencedate[dataset$occurrencedayofweek == ""])
week_day
dataset$occurrencedayofweek[(dataset$occurrencedayofweek == "")] <-strftime(substr(week_day,1,10), format = "%A")
dataset$occurrencedayofweek[(dataset$occurrencedayofweek == "Friday    ")]<-"Friday"
dataset$occurrencedayofweek[(dataset$occurrencedayofweek == "Monday    ")]<-"Monday"
dataset$occurrencedayofweek[(dataset$occurrencedayofweek == "Saturday  ")]<-"Saturday"
dataset$occurrencedayofweek[(dataset$occurrencedayofweek == "Sunday    ")]<-"Sunday"
dataset$occurrencedayofweek[(dataset$occurrencedayofweek == "Thursday  ")]<-"Thursday"
dataset$occurrencedayofweek[(dataset$occurrencedayofweek == "Tuesday   ")]<-"Tuesday"
dataset$occurrencedayofweek[(dataset$occurrencedayofweek == "Wednesday ")]<-"Wednesday"
levels(factor(dataset$occurrencedayofweek))
```

insights
Most frequent crime of all times
Match works as the map function it collects the rows of the unique categories in the given column
tabulate then counts the number of rows for each category 
which.max gets the max category
```{r}
  uniqx <- unique(na.omit(dataset$mci_category))
  uniqx[which.max(tabulate(match(dataset$mci_category, uniqx)))]
```
 
 The graph shows that assault is the most frequent type of crime of all times
 
```{r}
uniqx
counts<-tabulate(match(dataset$mci_category, uniqx))
barplot(counts, main = "Total crimes for each category", names.arg = uniqx, 
                          xlab = "Category type", ylab = "Total number", 
                          col = "red")
```
Total crimes for different categories of Theft Over as it is the least frequent crime type

```{r}
theft_over_rows<-dataset[dataset$mci_category == 'Theft Over',]
theft_over_offence<-unique(na.omit(theft_over_rows$offence))
theft_over_offence[2]
counts<-tabulate(match(theft_over_rows$offence, theft_over_offence))
#par(mar = c(10, 4, 10, 1)) 
barplot(counts, main = "Total for different categories of Theft Over ", names.arg = theft_over_offence, xlab = "Category type", ylab = "Total number", col = "blue", las = 2)
library(ggplot2)
library(scales)
df<-data.frame(counts,theft_over_offence)
p<-ggplot(data=df, aes(x=theft_over_offence, y=counts,fill=counts))+geom_bar(stat="identity")  +scale_x_discrete(labels = wrap_format(10))+geom_text(aes(label = counts), vjust = 0)+ggtitle(label="Total for different categories of Theft Over")+xlab(label="Category Type")+ylab("count")
p
```
Most frequent Category of Auto theft
```{r}
theft_over_rows<-dataset[dataset$mci_category == 'Auto Theft',]
theft_over_offence<-unique(na.omit(theft_over_rows$offence))
theft_over_offence
```
Assault insights
Total crimes for different categories of Assault as it is the most frequent category

```{r}
theft_over_rows<-dataset[dataset$mci_category == 'Assault',]
theft_over_offence<-unique(na.omit(theft_over_rows$offence))
theft_over_offence
counts<-tabulate(match(theft_over_rows$offence, theft_over_offence))
#par(mar = c(10, 4, 10, 1)) 
barplot(counts, main = "Total for different categories of Assault", names.arg = theft_over_offence, 
                          xlab = "Category type", ylab = "Total number", 
                          col = "blue", las = 2)

library(ggplot2)
library(scales)
df <- data.frame(counts, theft_over_offence)
options(repr.plot.width = 500, repr.plot.height =3)
theme_update(plot.title = element_text(hjust = 0.5))
p<-ggplot(data=df, aes(x=theft_over_offence, y=counts,fill=counts))+geom_bar(stat="identity")  +scale_x_discrete(labels = wrap_format(5))+geom_text(aes(label = counts), vjust = 0)+ggtitle(label="Total for different categories of Assault")+xlab(label="Category Type")+ylab("count")

p
```

Neighbourhood insights

Most dangerous Neighborhood analysis
Waterfront Communities-The Island is the most unsafe place
```{r}
  Neighbour <- unique(na.omit(dataset$Neighbourhood))

  most_dangerous_n<-Neighbour[which.max(tabulate(match(dataset$Neighbourhood, Neighbour)))]
  most_dangerous_n
```

Waterfront Communities-The Island insights
Assault is the most frequent type of crime in the most dangerous neighborhood
```{r}
neighbor_analysis<-dataset[dataset$Neighbourhood == most_dangerous_n,]
neighbor_crimes<-unique(na.omit(neighbor_analysis$mci_category))
counts<-tabulate(match(neighbor_analysis$mci_category, neighbor_crimes))
pie(c(counts),labels=paste(neighbor_crimes,round(100*counts/sum(counts),0),"%"),col=c("blue","red","yellow","green","pink","magenta","gray"),main=paste("Crime Analysis",most_dangerous_n)) 
```
Least Dangerous neighborhood analysis
```{r}
 least_dangerous_n<-Neighbour[which.min(tabulate(match(dataset$Neighbourhood, Neighbour)))]
least_dangerous_n
```

```{r}
neighbor_analysis<-dataset[dataset$Neighbourhood == least_dangerous_n,]
neighbor_crimes<-unique(na.omit(neighbor_analysis$mci_category))
counts<-tabulate(match(neighbor_analysis$mci_category, neighbor_crimes))
typeof(counts)
pie(c(counts),labels=paste(neighbor_crimes,round(100*counts/sum(counts),0),"%"),col=c("blue","red","yellow","green","pink","magenta","gray"),main=paste("Crime Analysis",least_dangerous_n)) 
```
Crime Rate over Years 1966-2022
As the graph shows the crime rate kept increasing from 2014 till 2019 
2019 was the peak
then the rate dropped until now, this means that the police station implemented better security precautions
```{r}
years <- unique(na.omit(dataset$occurrenceyear))
years[which.max(tabulate(match(dataset$occurrenceyear, years)))]

count<-tabulate(match(dataset$occurrenceyear, years))


library(ggplot2)
#ggplot(mapping= aes(x=months, y=count, group=0)) +geom_line()+geom_point()+geom_line()+ggtitle("Crimes across all months")
# Second variable

val<-data.frame(years,count)
val
theme_update(plot.title = element_text(hjust = 0.5))
ggplot(data=val, aes(x=years, y=count, group=2)) +geom_line()+ggtitle(label="Crime Rates over the years\n 1966-2022")
```
Crime rate analysis of months
```{r}
months <- unique(na.omit(dataset$occurrencemonth))
months

count<-tabulate(match(dataset$occurrencemonth, months))
count
#plot(y=count,type = "o", col = "red", xlab = "Years", ylab = "Rain fall", main = "Crimes rate across months of all years", x=months,xlim=c(1,12))

library(ggplot2)
#ggplot(mapping= aes(x=months, y=count, group=0)) +geom_line()+geom_point()+geom_line()+ggtitle("Crimes across all months")
# Second variable
levels(factor(dataset$occurrencemonth))

val<-data.frame(months,count)
theme_update(plot.title = element_text(hjust = 0.5))
ggplot(data=val, aes(x=months, y=count, group=1)) +geom_line()+
  geom_point()+xlim(1, 12)+scale_x_discrete(limits=(1:12))+ggtitle(label="Total number of crimes per month")
```

Time difference between occurence and reporting
```{r}

s <-aggregate(x= difftime(dataset$reporteddate, dataset$occurrencedate, units = "days") ,by = list(dataset$mci_categor),FUN = mean,na.rm=TRUE)

mean_time_difference<-data.frame(s)
mean_time_difference
mean_num<-as.numeric(mean_time_difference$x)
pie(c(mean_num),labels=paste(mean_time_difference$Group.1,round((mean_num),0),"days"),col=c("blue","red","yellow","green","pink","magenta","gray"),main=paste("Time difference between occurence and reporting")) 
```
insights on premises
We calculated the most frequent premises type in which each category of crime occurs
```{r}

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

t<-aggregate(x=dataset$premises_type  ,by = list(dataset$mci_category),FUN = Mode)
t

```
Day of the week on which the most number of crimes occur
we calculated the most frequent day on which each type of crime occured
we also represented the frequency of crimes per days of the week using a bar chart
```{r}

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
day <- unique(na.omit(dataset$occurrencedayofweek))
day[1]
count<-tabulate(match(dataset$occurrencedayofweek, day))
count
library(ggplot2)
df<-data.frame(day,count)
theme_update(plot.title = element_text(hjust = 0.5))
p<-ggplot(data=df, aes(x=day, y=count,fill=count))+geom_bar(stat="identity")  +geom_text(aes(label = count), vjust = 0)+ggtitle(label="Total number of crimes per days of the week")+xlab(label="days")+ylab("count")
p
days<-aggregate(x=dataset$occurrencedayofweek  ,by = list(dataset$mci_category),FUN = Mode)
days

```

