---
title: "Project"
output: html_notebook
---



```{r}
rm(list=ls())
ls()

setwd(getwd()) 
```

load csv
```{r}
dataset <- read.csv("Major_Crime_Indicators.csv")

```

insights
Most frequent crime of all times
Match works as the map function it collects the rows of the unique categories in the given column
tabulate then counts the number of rows for each category 
which.max gets the max category
```{r}
  uniqx <- unique(na.omit(dataset$mci_category))
  uniqx[which.max(tabulate(match(dataset$mci_category, uniqx)))]
```
 
 The graph shows that assault is the most frequent type of crime of all times
 
```{r}
uniqx
counts<-tabulate(match(dataset$mci_category, uniqx))
barplot(counts, main = "Total crimes for each category", names.arg = uniqx, 
                          xlab = "Category type", ylab = "Total number", 
                          col = "red")
```
```{r}
theft_over_rows<-dataset[dataset$mci_category == 'Theft Over',]
theft_over_offence<-unique(na.omit(theft_over_rows$offence))
theft_over_offence[2]
counts<-tabulate(match(theft_over_rows$offence, theft_over_offence))
#par(mar = c(10, 4, 10, 1)) 
barplot(counts, main = "Total for different categories of Theft Over ", names.arg = theft_over_offence, xlab = "Category type", ylab = "Total number", col = "blue", las = 2)
df<-data.frame(counts,theft_over_offence)
p<-ggplot(data=df, aes(x=theft_over_offence, y=counts,fill=counts))+geom_bar(stat="identity")  +scale_x_discrete(labels = wrap_format(10))+geom_text(aes(label = counts), vjust = 0)
p
```
```{r}
theft_over_rows<-dataset[dataset$mci_category == 'Auto Theft',]
theft_over_offence<-unique(na.omit(theft_over_rows$offence))
theft_over_offence
```
Assault insights

```{r}
theft_over_rows<-dataset[dataset$mci_category == 'Assault',]
theft_over_offence<-unique(na.omit(theft_over_rows$offence))
theft_over_offence[2]
counts<-tabulate(match(theft_over_rows$offence, theft_over_offence))
#par(mar = c(10, 4, 10, 1)) 
barplot(counts, main = "Total for different categories of Assault", names.arg = theft_over_offence, 
                          xlab = "Category type", ylab = "Total number", 
                          col = "blue", las = 2)

library(ggplot2)
library(scales)
df <- data.frame(counts, theft_over_offence)
options(repr.plot.width = 500, repr.plot.height =3)
p<-ggplot(data=df, aes(x=theft_over_offence, y=counts,fill=counts))+geom_bar(stat="identity")  +scale_x_discrete(labels = wrap_format(5))+geom_text(aes(label = counts), vjust = 0)
p
```

Neighbourhood insights

Most dangerous Neighborhood analysis
Waterfront Communities-The Island is the most unsafe place
```{r}
  Neighbour <- unique(na.omit(dataset$Neighbourhood))

  most_dangerous_n<-Neighbour[which.max(tabulate(match(dataset$Neighbourhood, Neighbour)))]
```

Waterfront Communities-The Island insights
Assault is the most frequent type of crime in the most dangerous neighborhood
```{r}
neighbor_analysis<-dataset[dataset$Neighbourhood == most_dangerous_n,]
neighbor_crimes<-unique(na.omit(neighbor_analysis$mci_category))
counts<-tabulate(match(neighbor_analysis$mci_category, neighbor_crimes))
pie(c(counts),labels=paste(neighbor_crimes,round(100*counts/sum(counts),0),"%"),col=c("blue","red","yellow","green","pink","magenta","gray"),main=paste("Crime Analysis",most_dangerous_n)) 
```
Least Dangerous neighborhood analysis
```{r}
 least_dangerous_n<-Neighbour[which.min(tabulate(match(dataset$Neighbourhood, Neighbour)))]
neighbor_analysis<-dataset[dataset$Neighbourhood == least_dangerous_n,]
neighbor_crimes<-unique(na.omit(neighbor_analysis$mci_category))
counts<-tabulate(match(neighbor_analysis$mci_category, neighbor_crimes))
typeof(counts)
pie(c(counts),labels=paste(neighbor_crimes,round(100*counts/sum(counts),0),"%"),col=c("blue","red","yellow","green","pink","magenta","gray"),main=paste("Crime Analysis",least_dangerous_n)) 
```
Crime Rate over Years from 2014-2022
As the graph shows the crime rate kept increasing from 2014 till 2019 
2019 was the peak
then the rate dropped until now, this means that the police station implemented better security precautions
```{r}
years <- unique(na.omit(dataset$reportedyear))
years[which.max(tabulate(match(dataset$reportedyear, years)))]

count<-tabulate(match(dataset$reportedyear, years))
plot(count,type = "o", col = "red", xlab = "Years", ylab = "Crime Number",
     main = "Crimes rate across the years", x=years)
```
Crime rate analysis of months
```{r}
months <- unique(na.omit(dataset$reportedmonth))
months

count<-tabulate(match(dataset$reportedmonth, months))
count
#plot(y=count,type = "o", col = "red", xlab = "Years", ylab = "Rain fall", main = "Crimes rate across months of all years", x=months,xlim=c(1,12))

library(ggplot2)
#ggplot(mapping= aes(x=months, y=count, group=0)) +geom_line()+geom_point()+geom_line()+ggtitle("Crimes across all months")
# Second variable
plot(y=count,type = "o", col = "red", xlab = "Months", ylab = "Crime Number",
     main = "Crimes rate across the years", x=c(1,2,3,4,5,6,7,8,9,10,11,12))

# #Now the p contains a plot 

```

